Collaboration

### 2011-02-25 workflow.Rnw
Changed main document to Sweave.

\documentclass[11pt]{article}
\usepackage{fancyvrb,natbib,url,array,dcolumn}

@@

\newcolumntype{.}{D{.}{.}{2.1}} 
\begin{document}
\maketitle

@@

An idea is born in a coffee shop, a seminar, a quiet walk in 2011. This idea inspires a seminar paper in 2011. And a conference paper arises from the seminar paper in 2012. A dissertation chapter descends from the conference paper in 2013. Other dissertation chapters take up 2014. A submission to a journal happens in 2015. Revision and resubmission wait until 2017 while harried editors, reviewers and author strive to balance teaching, service, and life. In 2018 a child is born and a paper is published. In 2020 a graduate student in a coffee shop reads the paper and has an idea that challenges the analysis in the paper. What would happen if only the author had thought to control for X? Or run one more MCMC chain? Or chosen a different likelihood function? Will the United Nations (now eager to act based the paper) make a wrong move? Time to go back to the analyses run sometime between 2011 and 2020 to check. Easy. Right? Collaboration between the past self and the present self should be simple --- after all you are the same person in 2011 and 2020, right? And the way Microsoft Word/Stata/SPSS/R/LISREL stores files and the way that your Windows/Mac machine reads and writes them is the same, right? And you know exactly which bit of code produced which table and which figure, right? And you wrote your code following Nagler's Maxims \citet{nagler1995coding} and being following King's Replication Standard \citet{king1995replication} right?

@@

The principle regarding data is to know where the data came from and what operations were performed on which set of data. I have seen command files which do not begin with a call to load data and I have despaired. What is the meaning of \Verb+please-fit var1 with var2+ if I do not know where var1 var2 come from.\footnote{The command \Verb+please-fit+ comes from the MayIPleaseDoStatistics package for R which emphasizes politeness in data analysis. It is highly recommended for use with toddlers.}

@@

Using R, for example: I would write

\begin{verbatim}
pdf('fig1.pdf')
plot(stuff)
dev.off()
\end{verbatim}

@@

\begin{Verbatim}
 <<fig1,fig=TRUE>>==
 plot(stuff)
 @
\end{Verbatim}

@@

<<maketable, echo=FALSE, results=tex>>=
##source("http://artsci.wustl.edu/~jgill/Models/graph.summary.R")
dem.nations.df<-read.csv("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv")

lm1<-lm(protac2000~Gini2004+meanpr,data=dem.nations.df)

@@

colnames(themat)[3:4]<-c("~{}","~{}")
thextab<-xtable(themat,align=c("l",rep(".",4)),
                caption="People living in countries with unequal
  income distributions report less protest activity to World Values
  Survey interviewers than people living in countries with relatively
  more equal income distributions, controlling for mean political
  rights as measured by Freedom House 1980--2000",label="tab:protest")

print(thextab,include.colnames=FALSE,
      add.to.row=list(pos=list(0,3),
                command=c("& \\multicolumn{1}{c}{Coef} & \\multicolumn{1}{c}{Std. Err.} 
& \\multicolumn{2}{c}{95\\% CI} \\\\", 
                  paste("\\multicolumn{4}{l}{n: ",n,",   $\\hat{\\sigma}$:",est.sigma,"} \\\\ "))
                )
      )
@

@@

Group work requires version control.  If you are using some kind of weak version control system (which is much much better than none at all), like Dropbox, then you must communicate with other folks in your group before you edit existing files. That is, only one of you can edit a given file at a time. This prevents your work (or your colleagues work) from getting lost when you both try to save the same file on top of each other. If you find that you are needing to work on the same files at the same time, then you should work on establishing your own shared version control system. Free options include launchpad, github, sourceforge for open source projects (i.e. papers you are writing which you are happy to share with others). Each of those services include paid versions too. One may also use Dropbox as a kind of server for version control: checking out files from the Dropbox directory into a local working directory.

@@

An excellent, simple, and robust version control system is to merely rename your files with the date and time of saving them:  thedoc.txt becomes thedoc25092009-22:40.txt. And then when you reach certain milestones you can rename the file accordingly: thedocAPSA2009.txt [for the one sent to discussants at APSA] or thedocAPSR2015.txt [for the version eventually sent to the APSR six years after you presented it at APSA]. The systems I mentioned above all allow this kind of thing and are more elegant, but you can do it by hand too as long as you don't mind taking up a lot of disk space and having many many ``thedoc...'' files around.

\section{Obey Nagler's Maxims}
Read \citet{nagler1995coding}. Comment your code. Luckily, if you are
using a literate programming practice (Sweave, odfWeave, using R2HTML,
etc..), you can write paragraphs to surround your code. If not, then
use whatever commenting system exists in your analysis system.

@@

\begin{verbatim}
<<reg1>>
##Fitting the regression with x/1000 for ease of interpretation
thelm<-lm(y~I(x/1000))
##Fits using a quadratic term as suggested by Blah(1965) did not enhance the fit of the model
##thelmsq<-lm(y~I(x/1000)^2)

@
\end{verbatim}

@@

(and read Jonathan Nagler's piece on this too)

@@

(6) Name your files with evocative and descriptive names. Your collaborators are less likely to call you at midnight asking for help if your files are named ``regressions.R'' than if your files are called ``temp'' (note the use of the extension .R to tell us that the file contains R commands. use extensions like this as a standard practice to help you and your computer get along.)

@@

>>>>

\documentclass[12pt]{article}
\usepackage{fancyvrb,natbib,url,array,dcolumn,float,booktabs}

@@

\newcolumntype{.}{D{.}{.}{1.2}}

\usepackage[noae,nogin]{Sweave} %Declaring this explicitly so that I can modify the Sweave environment here rather in Sweave.sty which will be c
\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=.5em,fontsize=\footnotesize} 
\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=.5em,fontsize=\footnotesize} 
\DefineVerbatimEnvironment{Scode}{Verbatim}{xleftmargin=.5em,fontsize=\footnotesize} 
\fvset{listparameters={\setlength{\topsep}{0pt}}} 
\renewenvironment{Schunk}{\vspace{\topsep}}{\vspace{\topsep}} 
\SweaveOpts{keep.source=TRUE}  %Show comments
\fvset{fontsize=\normalsize} %slightly smaller verbatim font

%% These next lines tell latex that it is ok to have a single graphic
%% taking up most of a page, and they also decrease the space arou
%% figures and tables.
\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.1}   
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}
\setlength{\intextsep}{2ex}
\setlength{\floatsep}{2ex}
\setlength{\textfloatsep}{2ex}

\begin{document}
\maketitle

<<initialize,echo=FALSE,results=hide>>= 
##First, just setup the \R environment:

options(SweaveHooks=list(
          fig=function(){
            par(mar=c(3.5, 3, 1.1, 0),
                pty="s",
                mgp=c(1.5,0.5,0),
                oma=c(0,0,0,0))},#,
          echo=function(){options(continue=" ") ##Don't show "+" prompts,
                         options(prompt=" ")}
          ),
        digits=3,
        scipen=8,
        width=100
        )

@

@@

An idea is born in a coffee shop, a seminar, a quiet walk. It is winter 2011. This idea inspires a seminar paper in the spring. And a conference paper arises from the seminar paper in 2012. A dissertation chapter descends from the conference paper in 2013. Other dissertation chapters take up 2014. A submission to a journal happens in 2015. Revision and resubmission wait until 2017 while harried editors, reviewers and author strive to balance teaching, service, and life. In 2018 a child is born and a paper is published. In 2020 a graduate student in a coffee shop reads the paper and has an idea that challenges the analysis in the paper. What would happen if only the author had thought to control for X? Or run one more MCMC chain? Or chosen a different likelihood function? Will the United Nations (now eager to act based the paper) make a wrong move? Time to go back to the analyses run sometime between 2011 and 2020 to check. Easy. Right? Collaboration between the past self and the present self should be simple --- after all you are the same person in 2011 and 2020, right? And the way Microsoft Word/Stata/SPSS/R/LISREL stores files and the way that your Windows/Mac machine reads and writes them is the same, right? And you know exactly which bit of code produced which table and which figure, right? And you wrote your code following Nagler's Maxims \citet{nagler1995coding} and being following King's Replication Standard \citet{king1995replication} right?

@@

The principle regarding data is to know where the data came from and what operations were performed on which set of data. I have seen command files which do not begin with a call to load data and I have despaired. What is the meaning of \Verb+please-fit(var1 with var2)+ if I do not know where var1 var2 come from.\footnote{The command \Verb+please-fit+ comes from the MayIPleaseDoStatistics package for R which emphasizes politeness in data analysis. It is highly recommended for use with toddlers.}

@@

Using R, for example, I might specify that the file \Verb+fig1.pdf+ was produced by the command, where the \Verb+please-plot+ function was loaded elsewhere.

\begin{verbatim}
pdf('fig1.pdf')
please-plot(outcome by explanatory)
dev.off()
\end{verbatim}

@@

<<read.data>>=

##This read.csv line is slightly more complex than needed because of
##the particular encoding used for characters in the source file.

dem.nations.df<-read.csv(url("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv",encoding="ISO_8859-1"),as.is=TRUE)

row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful

good.df<-na.omit(dem.nations.df[,c("Gini2004","protac2000","Nation","meanpr")])
@

<<fig1code,echo=TRUE,fig=FALSE,width=3,height=3>>=
##The code to produce the following figure
par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
with(good.df,plot(Gini2004/100,protac2000,
     xlab='Gini Coefficient 2004 (UNDP)',
     ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
     cex=.8))

with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
     text(Gini2004/100,protac2000,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
@

\begin{figure}[ht]
  \begin{center}
<<fig1plot,echo=FALSE,fig=TRUE,width=3,height=3>>= 
<<fig1code>>
@
    \caption{Protest activity by income inequality from \citep{norris2009data}.}\label{fig:giniprot}
  \end{center}
\end{figure}

@@

<<maketable, echo=FALSE, results=tex>>=
##source("http://artsci.wustl.edu/~jgill/Models/graph.summary.R")

lm1<-lm(protac2000~I(Gini2004/100)+meanpr,data=good.df)

@@

colnames(themat)[3:4]<-c("~{}","~{}")
row.names(themat)<-c("Intercept","Income Inequality (lower=more equal)","Mean Political Rights (lower=more rights)")
thextab<-xtable(themat,align=c("l",rep(".",4)),
                caption="People living in countries with unequal
  income distributions report less protest activity to World Values
  Survey interviewers than people living in countries with relatively
  more equal income distributions, adjusting for average political
  rights as measured by Freedom House 1980--2000. Data from \\citep{norris2009data}.",label="tab:protest")

print(thextab,include.colnames=FALSE,
      hline.after=NULL,
      add.to.row=list(pos=list(0,3),
                command=c("\\toprule \n & \\multicolumn{1}{c}{Coef} & \\multicolumn{1}{c}{Std. Err.} 
& \\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ", 
                  paste("& \\multicolumn{4}{c}{n: ",n,",   $\\hat{\\sigma}$:",est.sigma,"} \\\\ \\bottomrule"))
                )
      )
@

@@

Group work requires version control.\footnote{See X for more
  discussion of what version control is.} Many people are familiar
with ``Track Changes'' feature in OpenOffice/Word/Pages or the fact that
Dropbox allows one to recover previous versions of files. These are
both kinds of version control. When collaborating with yourself or
others, it is useful to see what has changed, to feel free to
experiment and then to dump the experiment in favor of previous work,
to have multiple ``releases'' of the same document (one to MPSA, one
to APSR, one to your parents) without requiring that spawn many copies
of the same document and risk confusion and clobbering. Clobbering is
what happens when your future self or your other collaborators saves
an old version of a file over a new version --- erasing good work by
accident.

Of course if you rely on Dropbox or ``track changes'' for version
control, you must communicate with other folks in your group before
you edit existing files. Only one of you can edit and save a given
file at a time. This prevents your work (or your colleagues work) from
getting lost when you both try to save the same file on top of each
other. If you find that you are needing to work on the same files at
the same time, then you should work on establishing your own shared
version control system. Free options include launchpad, github,
sourceforge for open source projects (i.e. papers you are writing
which you are happy to share with others). Each of those services
include paid versions too. One may also use Dropbox as a kind of
server for version control: checking out files from the Dropbox
directory into a local working directory. (Notice that this is
different from directly working on files within your Dropbox managed
directories.)

@@

Fancy version control systems are not required, however. I suspect
that Google Docs allows a kind of version tracking and collaboration
as well.  An excellent, simple, and robust version control system is
to merely rename your files with the date and time of saving them:
thedoc.tex becomes thedoc25092009-22:40.tex.  If you are wise enough to have saved your documents as plain text (Such as LaTeX source (with or without R/Stata code chunks) \footnote{A quick Google search of ``Sweave for Stata'' turned up lots of resources for literate programming with Stata.} then you can easily compare documents using the many utilities available for comparing text files [FileMerge on OS X is pretty, but ediff and diff for Unix are very useful, I am sure that Windows has many other options]. Adobe Acrobat allows one to compare differences in pdf files. OpenOffice supports a ``Compare Documents'' option.

When you reach certain milestones you can rename the file accordingly:
thedocAPSA2009.tex [for the one sent to discussants at APSA] or
thedocAPSR2015.tex [for the version eventually sent to the APSR six
years after you presented it at APSA]. The systems I mentioned above
all allow this kind of thing and are much more elegant and capable,
but you can do it by hand too as long as you don't mind taking up a
lot of disk space and having many many ``thedoc...''  files around.

\section{Obey Nagler's Maxims}
Read \citet{nagler1995coding}. Comment your code. Luckily, if you are
using a literate programming practice (Sweave, odfWeave, using R2HTML,
etc..), you can write paragraphs to surround your code as well a
technical comments in the code itself. If you are not using a strictly
literate programming practice then use whatever commenting protocol
exists in your analysis language.

@@

\begin{Verbatim}
<<reg1>> =
##Fitting the regression with x/1000 for ease of interpretation
thelm<-lm(y~I(x/1000)+z,data=fabulous.df)
##Fits using a quadratic term as suggested by Blah(1865) did not enhance the fit of the model
##thelmsq<-lm(y~I(x/1000)^2+z,data=fabulous.df)
@
\end{Verbatim}

@@

@@

Name your files with evocative and descriptive names. Your
collaborators are less likely to call you at midnight asking for help
if your files are named ``regressions.R'' than if your files are
called ``temp9'' or ``supercalifragilisticexpialidocious'' (note the
use of the extension .R to tell us that the file contains R
commands. use extensions like this as a standard practice to help you
and your computer get along.)

@@

\section{Exemplars}

Lots of people are thinking about ``reproducible research'' and ``literate programming'' these days. Google those terms. I have experimented with three systems so far: (1) for one paper I simply included my Sweave document and data files into a compressed archive; (2) for another more computing intensive paper...
